---
title: "S&P 500 Stock Data Interactive Dashboard"
author: "Franklin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r load-libraries, echo = FALSE}
library(dplyr)
library(plotly)
library(DT)
library(lubridate)
library(here)
```

```{r load-data, echo = FALSE}
# Load the S&P 500 five-year stock data
# Make sure all_stocks_5yr.csv is in your working directory
here::i_am('Final_Project_3a.Rmd')
stocks <- read.csv(here::here("raw_data/all_stocks_5yr.csv"), stringsAsFactors = FALSE)

# Parse date column
stocks$date <- as.Date(stocks$date)

# Remove rows with missing close price
stocks <- stocks %>% filter(!is.na(close), !is.na(volume))
```

---

## Widget 1: Interactive Stock Price Explorer

Use the dropdown menu to select any S&P 500 stock and explore its closing price trend
over the five-year period (2013–2018). Hover over any point on the line to see the
exact date, closing price, and trading volume. You can also zoom and pan to inspect
specific time periods.

```{r widget1-interactive-line-chart, fig.width=10, fig.height=5, echo = FALSE}
library(htmlwidgets)
library(jsonlite)

# Get sorted list of all stock tickers
all_tickers <- sort(unique(stocks$Name))

# Pre-compute per-ticker data as a named list (serialized to JSON for JS consumption)
# Each element: list of date strings, close prices, and volumes
ticker_data <- lapply(all_tickers, function(tk) {
  df <- stocks %>%
    filter(Name == tk) %>%
    arrange(date)
  list(
    dates   = format(df$date, "%Y-%m-%d"),
    close   = round(df$close, 2),
    volume  = df$volume
  )
})
names(ticker_data) <- all_tickers

# Initial ticker to show on load
init_ticker <- all_tickers[1]
init_data   <- ticker_data[[init_ticker]]

# Build the base plotly figure with just the initial ticker's data
fig <- plot_ly(
  x          = init_data$dates,
  y          = init_data$close,
  customdata = init_data$volume,
  type       = "scatter",
  mode       = "lines",
  name       = init_ticker,
  line       = list(color = "#1f77b4", width = 1.8),
  hovertemplate = paste0(
    "<b>%{meta}</b><br>",
    "Date: %{x}<br>",
    "Closing Price: $%{y:.2f}<br>",
    "Volume: %{customdata:,} shares<extra></extra>"
  ),
  meta = init_ticker
) %>%
  layout(
    title = list(
      text = "S&P 500 Daily Closing Price (2013\u20132018)",
      font = list(size = 18, color = "#1a1a2e")
    ),
    xaxis = list(
      title       = "Date",
      showgrid    = TRUE,
      gridcolor   = "#e8e8e8",
      rangeslider = list(visible = TRUE, thickness = 0.06)
    ),
    yaxis = list(
      title      = "Closing Price (USD)",
      showgrid   = TRUE,
      gridcolor  = "#e8e8e8",
      tickprefix = "$"
    ),
    hovermode = "x unified",
    margin    = list(t = 80, b = 60),
    plot_bgcolor  = "#fafafa",
    paper_bgcolor = "#ffffff",
    font = list(family = "Georgia, serif", size = 13)
  )

# Inject the full ticker dataset and a <select> dropdown via onRender
# This avoids embedding 505 plotly traces; instead JS swaps data on selection
ticker_json <- toJSON(ticker_data, auto_unbox = TRUE)

fig <- fig %>%
  onRender(
    sprintf(
      "
      function(el, x) {
        var allData = %s;
        var tickers = Object.keys(allData).sort();

        // -- Build the dropdown UI above the chart --
        var container = document.createElement('div');
        container.style.cssText = 'margin-bottom:8px; font-family: Georgia, serif;';

        var label = document.createElement('label');
        label.textContent = 'Select Stock Ticker: ';
        label.style.cssText = 'font-size:14px; font-weight:bold; color:#2c3e50; margin-right:6px;';

        var sel = document.createElement('select');
        sel.style.cssText = 'font-size:13px; padding:4px 8px; border-radius:4px; border:1px solid #aaa; background:#f9f9f9; cursor:pointer;';

        tickers.forEach(function(tk) {
          var opt = document.createElement('option');
          opt.value = tk;
          opt.textContent = tk;
          if (tk === '%s') opt.selected = true;
          sel.appendChild(opt);
        });

        container.appendChild(label);
        container.appendChild(sel);
        el.parentNode.insertBefore(container, el);

        // -- Update chart when selection changes --
        sel.addEventListener('change', function() {
          var tk  = this.value;
          var d   = allData[tk];
          Plotly.restyle(el, {
            x:          [d.dates],
            y:          [d.close],
            customdata: [d.volume],
            meta:       [tk],
            name:       [tk]
          }, [0]);
          // reset range slider so the full series is shown
          Plotly.relayout(el, {'xaxis.autorange': true, 'yaxis.autorange': true});
        });
      }
      ",
      ticker_json,
      init_ticker
    )
  )

fig
```

**Key Takeaways:** The closing price trajectories across S&P 500 stocks vary substantially, 
with some stocks (such as those in the technology sector) exhibiting strong multi-year 
upward trends while others remain flat or decline over the same period. The range slider 
and zoom functionality allow users to examine short-term volatility alongside the broader 
five-year trend, revealing that periods of high daily price movement often coincide with 
elevated trading volume visible in the tooltip.

---

## Widget 2: S&P 500 Stock Performance Summary Table

The table below summarizes key performance statistics for each of the 505 S&P 500 stocks 
over the five-year period. Use the search box to filter by ticker symbol, and click any 
column header to sort. This allows quick identification of the top performers, most 
volatile stocks, and highest-volume names in the dataset.

```{r widget2-summary-table, echo = FALSE}
# Compute per-stock summary statistics
summary_stats <- stocks %>%
  group_by(Name) %>%
  arrange(date) %>%
  summarise(
    Start_Date        = format(min(date), "%Y-%m-%d"),
    End_Date          = format(max(date), "%Y-%m-%d"),
    Start_Price_USD   = round(first(close), 2),
    End_Price_USD     = round(last(close), 2),
    Total_Return_Pct  = round((last(close) - first(close)) / first(close) * 100, 2),
    Annualized_Vol_Pct = round(
      sd(diff(log(close)), na.rm = TRUE) * sqrt(252) * 100, 2
    ),
    Max_Close_USD     = round(max(close, na.rm = TRUE), 2),
    Min_Close_USD     = round(min(close, na.rm = TRUE), 2),
    Avg_Daily_Volume  = round(mean(volume, na.rm = TRUE) / 1e6, 2),
    Trading_Days      = n(),
    .groups = "drop"
  ) %>%
  rename(
    Ticker               = Name,
    `Start Date`         = Start_Date,
    `End Date`           = End_Date,
    `Start Price ($)`    = Start_Price_USD,
    `End Price ($)`      = End_Price_USD,
    `Total Return (%)`   = Total_Return_Pct,
    `Annualized Volatility (%)` = Annualized_Vol_Pct,
    `Max Close ($)`      = Max_Close_USD,
    `Min Close ($)`      = Min_Close_USD,
    `Avg Daily Volume (M shares)` = Avg_Daily_Volume,
    `Trading Days`       = Trading_Days
  )

# Render interactive DT table
datatable(
  summary_stats,
  rownames  = FALSE,
  filter    = "top",
  extensions = c("Buttons", "Scroller"),
  options = list(
    dom        = "Bfrtip",
    buttons    = c("csv", "excel"),
    pageLength = 20,
    scrollX    = TRUE,
    scrollY    = "420px",
    scroller   = TRUE,
    autoWidth  = TRUE,
    columnDefs = list(
      list(className = "dt-center", targets = "_all")
    ),
    initComplete = JS(
      "function(settings, json) {",
      "  $(this.api().table().header()).css({'background-color': '#2c3e50', 'color': '#ffffff'});",
      "}"
    )
  ),
  caption = htmltools::tags$caption(
    style = "caption-side: top; text-align: left; font-size: 14px; color: #555;",
    "Table 1: S&P 500 Five-Year Performance Summary (2013–2018). ",
    "All price figures are in USD. Annualized volatility is computed from daily log returns ",
    "scaled by sqrt(252)."
  )
) %>%
  formatStyle(
    "Total Return (%)",
    background = styleInterval(
      c(-50, 0, 100, 300),
      c("#f8d7da", "#ffe5d0", "#d4edda", "#a8d5b5", "#5aab7a")
    ),
    fontWeight = "bold"
  ) %>%
  formatStyle(
    "Annualized Volatility (%)",
    background = styleInterval(
      c(20, 40, 60),
      c("#d4edda", "#fff3cd", "#ffd4a3", "#f8d7da")
    )
  )
```

**Key Takeaways:** Across the 505 S&P 500 constituents, total returns over the five-year 
period range dramatically — from deep losses exceeding -80% to gains above 1,000% — 
illustrating the wide dispersion of individual stock performance even within a broadly 
rising market index. Stocks with higher annualized volatility (highlighted in orange/red) 
tend to be concentrated in smaller-cap or sector-specific names, and sorting by 
"Avg Daily Volume" quickly reveals the most liquid large-cap stocks that dominate 
index-level trading activity.
```
